### 2013-08-22 doc/dissertation/document/chapters/parallel_mc/parallel_mc.tex
pushing full draft of disseration
moving parallel metrics to an appendix

From the data we see that for all solvers, the parallel results agree
with the reference serial results within the solution tolerance for
all DOF in the system. In addition, we do not notice any systematic
differences in the results as a function of number of blocks, number
of sets, or amount of overlap. In addition, the parallel GMRES results
have maximum absolute differences of the same magnitude as all MCSA
results and a larger minimum absolute difference than all MCSA results
by three orders of magnitude. Based on these results, MCSA
parallelized with the MSOD algorithm, and in particular the
implementation used here, generates the correct results with respect
to the serial conjugate gradient reference computation.

@@

%%---------------------------------------------------------------------------%%
\section{Parallel Performance Metrics\ }
\label{sec:parallel_performance_metrics}

How effectively the presented algorithms parallelize MCSA will be
determined by several performance measures. For this work, these
measures will come directly from Keyes' primer on the scalability of
domain decomposed methods and specifically his discussion on how to
measure their parallel performance \cite{keyes_how_1999}. All of
these metrics are derived from two measured quantities: the number of
iterations required for convergence and the total wall time required
for convergence. Each of these quantities are then measured in varying
parameter studies by changing the number of processes in a simulation
and by varying the local or global problem size. Using the notation of
Keyes, we will define \textit{parallel efficiency} as $\eta(N,P)$
where the efficiency is a function of the global problem size, $N$,
and the number of processes in the calculation, $P$. There are several
types of efficiencies that will be useful in our analysis of
scalability with different definitions depending on the type of
parametric scaling study performed. For all efficiencies, a value of
100\% is considered perfect while a value of 0\% signals the worst
possible performance. Often, as was observed in this work, it is the
case that efficiencies above 100\% are computed in certain scaling
studies. For the data presented in the following sections, an
explanation of each observed case of super-unitary efficiencies will
be provided.

Of primary interest to the nuclear engineer is the \textit{strong
  scaling} performance of an algorithm
\cite{siegel_analysis_2012}. In this type of scaling study, the
global problem size is fixed and the local problem size changes with
the number of processes in the calculation. In this case, as $P$
increases, $N$ is fixed, thus reducing the local problem size and
performance by having a larger ratio of communication to computational
work. This type of scaling is important because typically the global
problem size is fixed due to the problem of interest. Consider the
desire for a high fidelity neutronics simulation of the entire core of
a nuclear reactor. In this case, the global problem is known from the
geometry definition including the number of fuel assemblies and the
geometry of those assemblies all the way down to the resolution of the
space, angle, and energy discretizations required in order to capture
physical phenomena of interest. Because the global problem size is
known a priori, strong scaling is an important measure of how useful a
solution technique is when larger and larger values of $P$ are used to
solve the problem. Ideally, the time to solution for the the problem
will decrease linearly as a function of $P$, permitting a faster time
to solution using the maximum available resources and thus making sure
all $P$ are positively affecting the runtime of the problem.

For strong scaling, the objective is to then decrease the runtime of a
fixed size global problem as a linear function of $P$. Based on this,
we can define the strong scaling \textit{absolute efficiency} as:
\begin{equation}
  \eta_{strong}(N,P) = \frac{1}{P} \frac{T(N,1)}{T(N,P)}\:,
  \label{eq:strong_scaling_absolute}
\end{equation}
where $T(N,P)$ is the wall time for a computation of global size $N$
using $P$ processes. Note here that this particular definition is
measured with respect to a serial computation. On leadership class
machines, it is often the case that a problem size large enough to
effectively execute on $O(10,000)$ cores or larger is significantly
larger than any problem that may be solved by a single core due to
memory restrictions. Therefore, we consider absolute scaling for a
base case where $Q$ processes are used instead of a serial
computation:
\begin{equation}
  \eta_{strong}(N,P|Q) = \frac{Q}{P} \frac{T(N,Q)}{T(N,P)}\:.
  \label{eq:strong_scaling_absolute_ref}
\end{equation}
The absolute efficiency measure presented here only considers the
total time to solution. Often, that total time to solution may be
changing for an iterative method due to the fact that the number of
iterations required to converge may be changing as a function of
$P$. This will often happen in cases where modifying $N$ changes the
spectral character of the problem, often making it stiffer and
therefore more difficult to solve when $N$ increases. Therefore we
also consider \textit{algorithmic efficiency} from Keyes' work which
considers this potential change in iteration count:
\begin{equation}
  \eta_{alg}(N,P|Q) = \frac{I(N,Q)}{I(N,P)}\:,
  \label{eq:algorithmic_efficiency}
\end{equation}
where $I(N,P)$ is the number of iterations required to converge a
problem of global size $N$ computed using $P$ processes. We can then
adjust the absolute efficiency using this new algorithmic efficiency
in order to compute the parallel scalability of a single iteration
such that algorithmic effects may be neglected. We define the
\textit{implementation efficiency} as:
\begin{equation}
  \eta_{impl}(N,P|Q) = \frac{\eta}{\eta_{alg}}\:,
  \label{eq:implementation_efficiency}
\end{equation}
using the absolute efficiency in the numerator, giving an effective
measure of scalability of a single iteration of the algorithm.

In addition to strong scaling, in certain cases it is useful to
consider the \textit{weak scaling} of an iterative method. In this
case, the ratio $N/P$ is fixed such that the local problem size
remains the same for all values of $P$. In this case, the objective is
to maintain a constant wall time while both $N$ and $P$ grow as a
fixed ratio. In this case, we define the absolute efficiency as:
\begin{equation}
  \eta_{weak}(M|N,P|Q) = \frac{T(M,Q)}{T(N,P)}\:,
  \label{eq:weak_scaling_absolute}
\end{equation}
subject to the constraint $M/Q = N/P$. The scaling is perfect if the
runtime is static for all $P$ when $N/P$ is fixed.

Often, weak scaling is of interest to the nuclear engineer due to the
fact that the computational resources available are fixed and may not
be able to contain the entire global problem of interest. In this
case, an algorithm with good weak scaling performance permits all of
the computational resources available to be used effectively to allow
for the highest fidelity problem possible to be solved. In addition,
for the resources available, the memory to processor ratio on a
distributed machine is often fixed, giving an exercise in fixed $N/P$
to solve the highest fidelity problem possible. Therefore, good weak
scaling should not be viewed as an enhancement of time to solution for
the engineer but rather permitting a higher fidelity problem to be
solved within a given time constraint. For problems where increasing
fidelity increases the stiffness of the system, there is often a
trade-off between the absolute efficiencies computed for a given
method and the degrading algorithmic efficiencies that arise when
increasing the fidelity. Both time and hardware constraints should be
considered when preparing a parallel calculation.

%%---------------------------------------------------------------------------%%
\section{Leadership-Class Parallel Scaling Studies\ }
\label{sec:leadership_scaling_studies}
Using the parallel performance metrics presented in the previous
section, we will next measure performance in various scaling studies
using an implementation of the parallel MCSA algorithm to assess its 
quality using the neutron diffusion problem presented in
Appendix~\ref{chap:diffusion_problem}. For these scaling studies,
large problems at high levels of concurrency will be solved at a
leadership-class computing facility. In this section, we will outline
the parallel machine on which the calculations were performed and then
present a series of numerical studies that identify both strengths and
weaknesses of the parallel algorithm. Both pure domain decomposition
and MSOD parallel performance will be explored and compared to
performance results for the same problems using a set of production
parallel Krylov solvers.

>>>>

From the data we see that for all solvers the parallel results agree
with the reference serial results within the solution tolerance for
all DOF in the system. In addition, we do not notice any systematic
differences in the results as a function of number of blocks, number
of sets, or amount of overlap. In addition, the parallel GMRES results
have maximum absolute differences of the same magnitude as all MCSA
results and a larger minimum absolute difference than all MCSA results
by three orders of magnitude. Based on these results, MCSA
parallelized with the MSOD algorithm, and in particular the
implementation used here, generates the correct results with respect
to the serial conjugate gradient reference computation.

@@

%%---------------------------------------------------------------------------%%
\section{Leadership-Class Parallel Scaling Studies\ }
\label{sec:leadership_scaling_studies}
Using the parallel performance metrics presented in
Appendix~\ref{chap:parallel_metrics}, we will next measure performance
in various scaling studies using an implementation of the parallel
MCSA algorithm to assess its quality using the neutron diffusion
problem presented in Appendix~\ref{chap:diffusion_problem}. For these
scaling studies, large problems at high levels of concurrency will be
solved at a leadership-class computing facility. In this section, we
will outline the parallel machine on which the calculations were
performed and then present a series of numerical studies that identify
both strengths and weaknesses of the parallel algorithm. Both pure
domain decomposition and MSOD parallel performance will be explored
and compared to performance results for the same problems using a set
of production parallel Krylov solvers.
